---
output:
  html_document: default
  pdf_document: default
---
# Real-estate Transcation Behavior in Sacramento
# *by Anthony Terrence O'Brien Villate MD*

========================================================

```{r, include=FALSE}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.

library(knitr)
library(ggplot2)
library(GGally)
library(scales)
library(memisc)
library(stringr)
library(dplyr)
library(tidyr)
library(raster)
library(maps)
library(mapdata)
library(devtools)
library(ggmap)
library(psych)
library(gridExtra)
```

```{r, include=FALSE}
# Load the Data
#getwd() # read the working directory
# setwd('~Desktop') To set the working directory place the location after 
# the tilde
df <- read.csv("C:/Users/Anthony O\'Brien/Dropbox/1. Organized/1. Courses,lectures etc/9 .Nanodegree DAND/Term 2/Projects/Project  6 Explore and Summarise Data/Sacramento.csv")
```

# Abstract
In this analysis we explore dataon real estate transactions in 2016 over a five 
day period. The key goal of this study was to determine what were the main 
factors in the dataset which drove the price of the real estate transactions.

# Introduction
SpatialKey is a private company which buids geospatial insturance technology 
that simplifies how insureers view risk by building catastrophe data, mapping, 
and analytics. As part of their training resources, Spatialkey provides a 
plethora of datasets with geospatial information, which can be downloaded from
this url: https://support.spatialkey.com/spatialkey-sample-csv-data/

This.csv file is the Sacramento real estate transcations (SRET) file provided by 
SpatialKey. The file is a list of 985 real estate transactions in the Sacramento
county reported over a five-day period in 2016, as reported by the
[Sacramento Bee](http://www.sacbee.com/). 

[*The Sacramento Bee*](https://en.wikipedia.org/wiki/The_Sacramento_Bee)
is the largest newspaper in Sacramento, the fifth largest newspaper in 
California, and the 27th largest paper in the USA. 

[Sacramento County](https://en.wikipedia.org/wiki/Sacramento_County,_California)
is a county of the U.S State of California, and according to the 2010 census
the population was 1,418,788. Also it contains the state capital of California,
[Sacramento](https://en.wikipedia.org/wiki/Sacramento,_California), 
which has an estimated popualtion of 493,025; making it the 6th largest 
city in California.

# Summary of the Data Set

The SRET dataset was chosen as it provides a rich range of datatypes,
in a tidy format, with the option for practicing multiple editing,
analytical and visual techniques. 

The datset itself contains the following variables:

1. street: The physical address of the property.
            Categorical variable, and it will be dropped due to redundancy.

2. city: The city of the property within the Sacramento county.
            Categorical variable.
        
3. zip: The USPS Zone Improvement Plan used in postal addresses in the USA.
            Categorical variable organized in a numeric format.
  
4. state: The state in which the poperty is located.
            Categorical variable, which will be dropped as it does not provide
            useful information. All proerties are in the state of California
            
5. beds: The number of beds the property has.
            Quantitative variable    
    
6. baths: The number of bathrooms the property has.
            Quantitative variable    
      
7. sq__ft: The square feet the property has.
            Quantitative variable which will be converted to square meters
            
8. type: The type of property. 
            Categorical variable

9. sale_date: The day the property was sold.
            timedate variable which will be parsed into day

10. price: The price at which the propery was sold.
            Quantitative variable    

11. latitude: The geospatial location based on latitude of the property.
            Geospatial data

12. longnitude: The geospatial location based on longnitute of the property.
            Geospatial data

*Please note, all referenced websites were viewed on April 18th 2018*

#Dataset exploration and transformation

First I review the dataset dimensions, variables and structure

```{r, include=FALSE}
dim(df) #to test the number of rows and columns in the dataset (i.e. dimensions)
names(df) # to test the header names of the dataser
str(df) # structure of the dataset
```

The above dataset has a few variables which I want to drop or change. 
Particularly I want to remove street, stat and sale_datee, and 
convert sq__ft to square meters.

```{r, include=TRUE}
df <- subset(df, select = -c(street, state))# drops street and state variable
df['sq_m'] <- df$sq__ft*0.093# add a new column for data which is square meters
#test <-strsplit(as.character(df$sale_date), 'May') #parse data
#df['sale_day'] <-do.call(rbind, test) #add the day to the datafram
#df <- subset(df, select = -c(sale_date)) #drop the datetime
df <- subset(df, select = -c(sq__ft)) #drop the square feet
```

I review the dataset after the transformation.   

```{r, include=FALSE}
dim(df) #to test the number of rows and columns in the dataset (i.e. dimensions)
names(df) # to test the header names of the dataset
str(df) # structure of the dataset
```

# Univariate Plots Section

## Summary of city data

As a quick overview of the cities included in the analysis, I perform a quick,
aggregate of the variable city using a table, and then by visualizing the data
in a bar and pie chart

```{r, include=FALSE}
summary(df$city)
```

```{r, include=FALSE}
ggplot(df, aes(x=city, fill=city)) + 
  geom_bar() #simple bar chart
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
bp<- ggplot(df, aes(x="", y=city, fill=city))+
geom_bar(width = 1, stat = "identity")
pie <- bp + coord_polar("y", start=0) + theme(axis.text = element_blank())
#convert barchart data into circular coordinates
pie
```

The pie chart code is derived from the following url http://www.sthda.com.

What stands out form this initial analysis is that there are 3 cities with
an extraordinarly higher number of transactions relative to the other cities.
These three cities are Sacramento, Elk Grove and Lincoln.

## Summary of house types

From this point onwards I perform similar exploratory analysis and
visualizations on the remainding variables. 

```{r, include=FALSE}
summary(df$type)
```

```{r,  echo=FALSE, message=FALSE, warning=FALSE}
ggplot(df, aes(x=type, fill=type)) + 
  geom_bar()
```

From initial inspection it does not seem that the variable type will provide 
much insight into the data. 

## Summary of number of beds

```{r, include=TRUE,}
summary(df$beds)
```

```{r, include=TRUE,}
ggplot(df, aes(x=beds, fill=beds)) + 
  geom_bar()
```

There is a condierable number of missing data in beds, which can be later 
removed. Also, This data seems like it would provide useful information
for understanding the transactional patterns in the data. Also at present 
this variable is an integer which should be converted to a factor 
variable to really appreciate its usefulness. 

## Summary of number of baths

```{r, include=TRUE,}
summary(df$baths)
```

```{r, include=TRUE,}
ggplot(df, aes(x=baths, fill=baths)) + 
  geom_bar()
```

The variable baths has similar properties as the variable beds. 

## Summary of sales price

```{r, include=TRUE,}
summary(df$price)
```

```{r, include=TRUE,}
ggplot(aes(x = price), data = df) +
  geom_histogram() +
  scale_x_continuous()
```

This is one of the most important variable sin this dataset as we are 
studying property sales. Visually we note a high number of missing
values and that the data is skewed to the right due to high sales numbers.

## Summary of square meters for property 

```{r, include=FALSE}
summary(df$sq_m)
```

```{r,  echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x = sq_m), data = df) +
  geom_histogram() +
  scale_x_continuous()
```

This variable shares similar proerties like the varibale price

# Univariate Analysis

With this initial univariate analysis, there are interesting findings to report:

1) There is a substantial amount of property located in Sacramento (n=434),
followed by Elk Grove (n=114) and Lincoln (n=72).

2) Most properties are reported as Residential.

3) Most properties have 3 bedrooms and 2 bathrooms.

4) The median cost of properties is $213,800, however there are also outlier
above \$750,000 with a max of $884,800.

5) The median square meter space of the properties is 121.3 m^2, with the third
quartile in 159.80m^2. Interestingly the max area reported is 541.4. 

6) In the square meter area for properties there is a significant amount of
properties that report 0 as their area. The same for price. 

Some initial questions which this analysis provokes are:

1) It would be interesting to see if there are any particular properties of
Sacramento, Elk Grove and Lincoln, relative to the available data,
that can suggest why they are so popular as a sales area.

2) Additionally it will be interesting to run similar analysis for properties 
which are considerably outside the normal ditribution of the data. For example,
this can include data for properties which are exceptionally pricey or that have
large surface area for sale. 

3) Likewise it will be interesting to see if there are any particular patterns 
for unreported data (eg. they may be from the same area, or for certain 
listings).

# Bivariate Plots and Analysis Section

The variables ''latitude' and 'longitude' contain geospatial data. This data
can be used to create a scatterplot map of the property areas. The goespatial
information can be used to visually observe sales patterns in the data. 
Potentially helping us to generate hypothesis to answer questions raised in
the Univariate analysis section.

##What is the geospatial relationship between sales transactions and location?

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x=longitude, y=latitude), data=df)+
  geom_point() #scatter plot 
```

Leveraging the geospatial data, we tag its color using the city variable

```{r,  echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x=longitude, y=latitude, color= city), data=df)+
  geom_point() #scatter plot colored by city
```

With the above information we can overlay a map to observe how prevalent
sales are relative to the city in which they were sold. 

Therefore, this data is combined with map images pulled from google maps

```{r,  echo=FALSE, message=FALSE, warning=FALSE}
bc_bbox <- make_bbox(lat = latitude, lon = longitude, data = df) #create a bounding box according to latitude and longitude
bc_bbox
bc_big <- get_map(location = bc_bbox, source = "google", maptype = "terrain") #call on google maps and bounding box to set map frame for scatter plot data
ggmap(bc_big) + 
  geom_point(data = df, mapping = aes(x = longitude, y = latitude, color= city)) #plot scatter plot on google maps
```

This mapping procedure was derived from teh following url:
http://eriqande.github.io/rep-res-web/lectures/making-maps-with-R.html

From this analysis we see that the three largest areas are representative of the
following: 1) Sacramento is Californias capital city and therefore a main point
of concentration for sales. 2) The next two main areas for sales (i.e. Elk 
Grove and Lincoln) are in the immediate vicinity of Sacramento (the south and 
North East, respectively). We also see that the further the areas are from 
Sacramento the less number of sales transcations occur. 

##Does the type of property affect the distribution of the sale?

```{r,  echo=FALSE, message=FALSE, warning=FALSE}
bc_bbox <- make_bbox(lat = latitude, lon = longitude, data = df)
bc_bbox
bc_big <- get_map(location = bc_bbox, source = "google", maptype = "terrain")
ggmap(bc_big) + 
  geom_point(data = df, mapping = aes(x = longitude, y = latitude, color= type))
```

Bsaed on the map image above it does not seem that the type of property affects
the geospatial distribution of the sale. It seems that the type of property is 
not a useful variable for describing the data. Frpm this point on I will not
include the type of property in the analysis.

## Distribution matrix to find hidden relationships

```{r, echo=FALSE, message=FALSE, warning=FALSE}
pairs.panels(df[,1:9], pch = 19)
```

From the  distribution matrix we see that:

1) The number of bedrooms and bathrooms a property has is highly correlated (i.e
as the number of bedrooms increases the number of bathrroms increases, r = 0.84)

2) Similarly, there is a weak positive relationship between the price of a 
property and the number of bedrooms and bathrooms (r = 0.34 and r = 0.42 
respectively).

3) Expectedly there is also a moderate posiitve relationshp between the 
square meter area and number of bedrooms (r=0.68) and bathrooms(r = 0.67) 

4) Intersetingly there is also a weak positive relationship between the price 
and the square meter area (r = 0.33). 

The above data suprisingly suggests that the price of a property while related
to its square meter area is not that substantially correlated to size. 

After reviewing the distribution matrix, I was curious if the location
may play more of a factor in determining sales prices rather than the 
number of bedrooms, bathrooms or size. 

## Converting integers to factors 

Knowing that the above variables for beds and baths are set as an integer, 
I convert them to factor variables and plot them in relation to the size
and price of the property sold. 

```{r, include=TRUE}
df[,'beds_factor'] <- as.factor(df[,'beds']) #convert data to factor variables
df[,'baths_factor'] <- as.factor(df[,'baths'])
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
p1 <- ggplot(df, aes(beds_factor, sq_m, color=beds_factor)) +
  geom_boxplot()
p2 <- ggplot(df, aes(baths_factor, sq_m, color=baths_factor)) +
  geom_boxplot()
p3 <- ggplot(df, aes(beds_factor, price, color=beds_factor)) +
  geom_boxplot()
p4 <-ggplot(df, aes(baths_factor, price, color=baths_factor)) +
  geom_boxplot()
grid.arrange(p1, p3, p2, p4)
```

The above grid method layout was derived from the following URL:
https://cran.r-project.org/web/packages/egg/vignettes/Ecosystem.html

At this point I replace 0 with NA and remove from the plot

```{r, include=TRUE}
df[, 11:12][df[, 11:12] == 0] <- NA #convert 0 to NA
```

And then I replot the above graphs excluding the missing variables

```{r,  echo=FALSE, message=FALSE, warning=FALSE}
p1 <- ggplot(df[!is.na(df$beds_factor), ], aes(beds_factor, sq_m, color=beds_factor)) +
  geom_boxplot()
p2 <- ggplot(df[!is.na(df$baths_factor), ], aes(baths_factor, sq_m, color=baths_factor)) +
  geom_boxplot()
p3 <- ggplot(df[!is.na(df$beds_factor), ], aes(beds_factor, price, color=beds_factor)) +
  geom_boxplot()
p4 <-ggplot(df[!is.na(df$baths_factor), ], aes(baths_factor, price, color=baths_factor)) +
  geom_boxplot()
grid.arrange(p1, p3, p2, p4)
```

The method of NA removal was obtained from the following URL:
https://stackoverflow.com/questions/11403104/remove-unused-factor-
levels-from-a-ggplot-bar-plot

By converting the number of bedrooms and the number of bathrooms to a factor 
variable, we unveil a relationship in which in the size per square meter and the
price increase relative to the number of bathrooms and bedrooms.

Altough it is intersting to see that this relationship does not exactly hold
for properties with 6 or more bedrooms or 5 bathrooms. Other aspects like
location may have influenced this outcome for price. Though it is not clear
why th size of location would decrease relative to the number of rooms.

Given the relationship of bedroom and bathroom with price and
size,  plot thevariable beds and baths relative to the geospatial
data. 

```{r,  echo=FALSE, message=FALSE, warning=FALSE}
bc_bbox <- make_bbox(lat = latitude, lon = longitude, data = df)
bc_bbox
bc_big <- get_map(location = bc_bbox, source = "google", maptype = "terrain")
ggmap(bc_big) + 
  geom_point(data = df, mapping = aes(x = longitude, y = latitude, color= beds_factor))
```

In general it is difficult to discern any patterns, other than areas further
away from Sacramento (South, South-East, North and North-East) have a 
higher number of bedrooms relative to Sacramento.

```{r,  echo=FALSE, message=FALSE, warning=FALSE}
bc_bbox <- make_bbox(lat = latitude, lon = longitude, data = df)
bc_bbox
bc_big <- get_map(location = bc_bbox, source = "google", maptype = "terrain")
ggmap(bc_big) + 
  geom_point(data = df, mapping = aes(x = longitude, y = latitude, color= baths_factor))
```

Like number of bedrooms, there is a distribution which seems to lead to
increased number of bathrooms the furhter away one mooves from Sacramento.

## Price range according to geographic location

```{r,  echo=FALSE, message=FALSE, warning=FALSE}
bc_bbox <- make_bbox(lat = latitude, lon = longitude, data = df)
bc_bbox
bc_big <- get_map(location = bc_bbox, source = "google", maptype = "terrain")
ggmap(bc_big) + 
  geom_point(data = df, mapping = aes(x = longitude, y = latitude, color= price))
```

The above map suggests that tmore expensive areas are found outside of the 
Sacramento area, and more towards the north east around Granite Bay, Folsom, 
El Dorado Hills, and Cameron Park. While a couple of more expensive sales also
occured in the bottom right corner of the map around the Wilton area. 

## Location and size

```{r,  echo=FALSE, message=FALSE, warning=FALSE}
bc_bbox <- make_bbox(lat = latitude, lon = longitude, data = df)
bc_bbox
bc_big <- get_map(location = bc_bbox, source = "google", maptype = "terrain")
ggmap(bc_big) + 
  geom_point(data = df, mapping = aes(x = longitude, y = latitude, color= sq_m))
```

While not as prominent a feature, as one goes out of Sacramento area the 
square meter area per property tends to increase. 

## Price and size

Prior analysis suggests that there is a linear relationship between
size and property. Therefore to study this relationship I perform a 
scatterplot.

```{r, include=FALSE}
w1 <- ggplot(aes(x = price, y = sq_m), data = df)+
  geom_point()
w2 <-ggplot(aes(x = price, y=sq_m, color=beds_factor), data=df)+
  geom_point()
w3 <-ggplot(aes(x=price, y=sq_m, color=baths_factor), data=df)+
  geom_point()
w4 <-ggplot(aes(x=price, y=sq_m, color=city), data=df)+
  geom_point()
grid.arrange(w1) #use of gird arrange to plot following plots
```

In general there is a visual relationship betwen price and size. However this 
relationship seems to hold until around $375,000.

To confirm this observation I plot a linera regression and lowess plot, on top
of the scatterplot. 

```{r,  echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x = price, y = sq_m), data = df) + 
  geom_point(alpha=1/6, color='BLACK') +
  geom_smooth(method = 'lm', se = TRUE, color = 'red') +
  geom_smooth(color = 'blue', se = FALSE) #simple scatter plot with regression line and loess line 
```

From here, I zoom in on to the data and then readjust the plot to redraw the 
regression lines as if the data after $375,000 was not considered

```{r,  echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x = price, y = sq_m), data = df) + 
  geom_point(alpha=1/10, color='BLACK') +
  geom_smooth(method = 'lm', se = FALSE, color = 'red') +
  geom_smooth(color = 'blue', se = TRUE) +
  coord_cartesian(xlim = c(1551, 375000), ylim = c(15, 225)) #zoom on x and y axis
```

```{r,  echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x = price, y = sq_m), data = df) + 
  geom_point(alpha=1/10, color='BLACK') +
  geom_smooth(method = 'lm', se = FALSE, color = 'red') +
  geom_smooth(color = 'blue', se = TRUE) +
  xlim(1551, 375000) +
  ylim(15, 225) #removal of x and y axis limits
```

We see that when we consider the parsed data alone that there is a strong linear
relationship between size and price.

At this point I hypothesize that the relationship between price and size is 
relative to the properties location. Similarly, as number of bedrooms and bath
rooms relate to size, I plotted a similar scatter Plot

```{r,  echo=FALSE, message=FALSE, warning=FALSE}
grid.arrange(w4)
```

```{r,  echo=FALSE, message=FALSE, warning=FALSE}
grid.arrange(w2, w3)
```

We see that most of the pricing data is determined by Sacramento, and that 
anything under or equall to the median number of bedrooms and bathrooms is 
generally agregated within the area represented by Sacramento.  


# Multivariate Plots and Analysis Section

Now that I have visually analyzed the data, I will run a few univariate linear 
regression models separately to determine which significant relationships cna be 
added to a multiple linear regression model. 

## Regresion model

```{r, include=FALSE,}
m1 <- lm(I(price) ~ I(sq_m), data=df)
summary(m1)
mtable(m1)
```

```{r,  echo=FALSE, message=FALSE, warning=FALSE}
m1 <- lm(I(price) ~ I(city), data=df)
summary(m1)
mtable(m1)
```

```{r,  echo=FALSE, message=FALSE, warning=FALSE}
m1 <- lm(I(price) ~ I(beds_factor), data=df)
summary(m1)
mtable(m1)
```

```{r,  echo=FALSE, message=FALSE, warning=FALSE}
m1 <- lm(I(price) ~ I(baths_factor), data=df)
summary(m1)
mtable(m1)
```

From the initial univariate analysis, we see that city relates best to the 
transaction listings, followed by number of bedrooms, then bathrooms, and 
finally square meters. 

Using this insight the multivariate model will be created accordingly.

```{r,  echo=FALSE, message=FALSE, warning=FALSE}
m1 <- lm(I(price) ~ I(city), data=df)
m2 <- update(m1, ~ . + beds_factor)
m3 <- update(m2, ~ . + baths_factor)
m4 <- update(m3, ~ . + sq_m)
mtable(m1, m2, m3, m4)
```

From the initial univariate regression analysis, we see that the most
expensive area is Granite Ville. However upon considering square meter, number
of bedrooms and bathrooms Garden Valley ends up having the most expensive
proeprties, then Granite Ville. 

# Visualization of sales

```{r,  echo=FALSE, message=FALSE, warning=FALSE}
bc_bbox <- make_bbox(lat = latitude, lon = longitude, data = df)
bc_big <- get_map(location = bc_bbox, source = "google", maptype = "roadmap", zoom=10)
a1 <- ggmap(bc_big) + 
  geom_point(data = df, mapping = aes(x = longitude, y = latitude, color=city, alpha=1/500))+
  stat_ellipse(data= df, aes(x=longitude, y=latitude, group=city),type = "norm")+
  labs(title="Geospatial distribution of sales by city", y="Latitude", x="Longitude")
p5 <- a1 + theme(legend.position="left") + theme(legend.position='none') + theme(plot.title = element_text(hjust = 0.5, size=11)) #the above is the same method for plotting maps except I also save them to a temporary file to manipulate hte titles and legends
```

```{r,  echo=FALSE, message=FALSE, warning=FALSE}
bc_bbox <- make_bbox(lat = latitude, lon = longitude, data = df)
bc_big <- get_map(location = bc_bbox, source = "google", maptype = "roadmap", zoom=10)
a2 <- ggmap(bc_big) + 
  geom_point(data = df, mapping = aes(x = longitude, y = latitude, color=price))+
  stat_ellipse(data= df, aes(x=longitude, y=latitude, color=price, group=city),type = "norm")+
  labs(title="Geospatial distribution of sales by price", y="Latitude", x="Longitude")
p6 <- a2 + theme(legend.position="left") + theme(plot.title = element_text(hjust = 0.5, size=11)) + labs(color='Price') + theme(legend.position='none')  

```

```{r,  echo=FALSE, message=FALSE, warning=FALSE}
grid.arrange(p5, p6, ncol=2, bottom="In the graph to the right, lighter colors are equivalent to higher sales")
```

From the above graph we note that the sales density is largest towards 
Sacramento and immediately adjascent areas. However, sales prices increase
the further out the property is relative to Sacramento.

This is interesting because generally in large cities in the Americas, property
prices tend to increaes the closer one is to the main city, for example this 
is observed in San Francisco, Boston, New York, and even cities outside the
USA like Bogota, Colombia. 

The pricing patterns are paradoxical to what one would traditionally expect. 

------

# Final Plots and Summary

> **Tip**: You've done a lot of exploration and have built up an understanding
of the structure of and relationships between the variables in your dataset.
Here, you will select three plots from all of your previous exploration to
present here as a summary of some of your most interesting findings. Make sure
that you have refined your selected plots for good titling, axis labels (with
units), and good aesthetic choices (e.g. color, transparency). After each plot,
make sure you justify why you chose each plot by describing what it shows.

### Plot One
```{r  echo=FALSE, message=FALSE, warning=FALSE}


d1 <-ggplot(aes(x=price, y=sq_m, color=city), data=df)+
  geom_point() +
  geom_smooth(method = 'lm', se = FALSE, color = 'red') +
  geom_smooth(color = 'blue', se = TRUE) + 
  labs(y="Size of property in square meters", x=" Cost of property in USD")+
  ggtitle("Cost of property relative to size and location") 



d3 <- ggplot(aes(x=price, y=sq_m, color=city), data=df)+
  geom_point() +
  geom_smooth(method = 'lm', se = FALSE, color = 'red') +
  geom_smooth(color = 'blue', se = TRUE) +
  xlim(1551, 375000) +
  ylim(15, 225) +
  labs(y="Size of property in square meters", x="Price in USD") + theme(legend.position="right")

 
d1 + theme(plot.title = element_text(hjust = 0.5, size=11))
```

### Description One

What we observe when plotting property prices relative to the location is that
Sacramento city plays a huge role in stabilizing pricing property. As the 
property gets further away from Sacramento, prices begin to become less linearly
related to the property's size. 


### Plot Two
```{r, echo=FALSE, message=FALSE, warning=FALSE}
bc_bbox <- make_bbox(lat = latitude, lon = longitude, data = df)
bc_big <- get_map(location = bc_bbox, source = "google", maptype = "terrain", zoom=10)
a2 <- ggmap(bc_big) + 
  geom_point(data = df, mapping = aes(x = longitude, y = latitude, color=price))+
  stat_ellipse(data= df, aes(x=longitude, y=latitude, color=price, group=city),type = "norm")+
  labs(title="Geospatial distribution of sales by price", y="Latitude", x="Longitude")
p6 <- a2 + theme(legend.position="right") + theme(plot.title = element_text(hjust = 0.5, size=11)) + labs(color='Price (USD)') 
p6
```

### Description Two

This plot helps to confirm insight provided by plot one, in which
we observe that prices tend to rise the further away one gets from Sacramento.

### Plot Three

```{r, echo=FALSE, message=FALSE, warning=FALSE}
p3 <- ggplot(df[!is.na(df$beds_factor), ], aes(beds_factor, price, color=beds_factor)) +
  geom_boxplot() +
  labs(y="Price of property in USD", x="Number of bedrooms") + theme(legend.position='none')  
p4 <-ggplot(df[!is.na(df$baths_factor), ], aes(baths_factor, price, color=baths_factor)) +
  geom_boxplot() + 
  labs(y="Price of property in USD", x="Number of bathrooms") + theme(legend.position='none')  
grid.arrange(p3, p4, top="Cost of propery relative to number of bedrooms or bathrooms")
```
```

### Description Three

In the third visualization, we note acknowledge there are other factors
which can influence the price of a transaction other than location and size,
such as the number of bedrooms and bathrooms. 

In fact, we also note this in our multi-linear regression model, and point out
how these two co-variates impact the cost of the most expensive property
in this sample (see multi-linear regression section).

------

# Reflection

I was interested in analyzing this particular set of data, because it contained
an appropriate number of geolocalization information to construct geospatial
maps, while still allowing me to apply the different techniques learnt during 
the exploratory data analysis course. In addition to this, I am currently
searching for property to rent and I thought it woudl be useful to leverage
this opportunity to gain insight into a real life scenario. 

A lot of the coding was straight forward, however I did have to try multiple
different tools/techniques before being able to graph the map relative to the
available data. I woudl say anything relating map data to the current data was
a bit of a challenge, however geospatial maps were acomplished. Another 
challenge I ran into was with parsing the date-time data. In the end, I dropped 
the data, however in my initial attempt, I was able to separate the day from 
the rest of the date-time data. Ulteimately the data was not useful so I dropped
it. 

I learnt to transform values in a column like 0 into NaN or NA, and recently 
learnt that NA is a non-numeric value, while NaN is a numeric value. 

Ultimately while I see the power of R to compute big data in mere seconds, and
perform complex visualizations. I still prefer python. I'm also quite satisfied
with the linear regression model generation. It was suprising to me that
location was more indicative of pricing when compared to over variables like 
size. I guess the old saying "location, location, location" really makes an 
impact afterall. In the future this data, can be combined with a yearly account
to really try and grasp deeper tendencies (afterall this data only represents
sales in a 5 day period). Additionally, other variables like family size 
would be interesting to integrate. Finally I would recommend that the creators
of the dataset reorganize the variable type, because at present it is not very
informative. 